Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties

### 2022-02-13 papers/paper/preprint.tex
Working on getting figure captions together

% authors
\usepackage{authblk}
\title{Which Scenario Should We Design For? Insights from House Elevation for the Multiple PDF Problem}
\author[1]{James Doss-Gollin}
\author[2]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{slr}{SLR}{relative sea level rise}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

\begin{table}
    \centering
    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name           & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}      & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge    & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
        Depth-damage   & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height & $h_0$            & Deterministic: \SI{0.5}{ft} above \gls{bfe} unless otherwise noted                   & TBD                   \\
        \bottomrule
    \end{tabular}
\end{table}

@@

\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    \includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:msl-boxplots}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[height=1.5in]{historic_surge.pdf}%
    \includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE).
        (a):
        Time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Blue line shows a maximum likelihood \gls{gev} fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\textwidth]{scenario_maps.png}
    \caption{
        Scenario maps showing the dependence of average lifetime cost as a function of \gls{slr} in 2100 for several values of $\Delta h$.
        One dot represents one \gls{sow}.
        %TODO
    }\label{fig:scenario-map}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
    \includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
    \caption{
        Need a summary sentence.
        %TODO
        Trade-offs between (a) construction cost and expected lifetime costs or (b) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
    }
    \label{fig:tradeoffs}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=3in]{lsl_priors.pdf}~
    \includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
    \caption{
        The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
        Alternative priors can provide robustness checks.
        Need to add panel labels, update data, \etc.
        %TODO
    }\label{fig:tradeoffs}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=6in]{inference_weights.pdf}
    \caption{
        The average weight assigned to \glspl{sow} from each model (\gls{rcp} scenario and ice sheet parameterization) for both priors considered.
        If the sampling distribution (green line) is used, then the weight assigned to each \gls{sow} is the same, regardless of whether it is over- or under-sampled.
    }\label{fig:model-average-weights}
\end{figure}

@@

\printbibliography
\end{document}

>>>>

% authors
\usepackage{authblk}
\title{Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties}
\author[1]{James Doss-Gollin}
\author[2]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym[]{cdf}{CDF}{cumulative distribution function}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{slr}{SLR}{relative sea level rise}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

\begin{table}
    \centering
    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \ref{sec:ead}         \\
        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 & TBD                   \\
        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                & TBD                   \\
        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   & TBD                   \\
        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       & TBD                   \\
        \bottomrule
    \end{tabular}
\end{table}

@@

\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    %\includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:msl-boxplots}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics[height=1.5in]{historic_surge.pdf}%
    %\includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE).
        (a):
        Time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Blue line shows a maximum likelihood \gls{gev} fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

@@

\begin{figure}
    %\includegraphics[width=\textwidth]{scenario_maps.png}
    \caption{
        Scenario maps showing the dependence of average lifetime cost as a function of \gls{slr} in 2100 for several values of $\Delta h$.
        One dot represents one \gls{sow}.
        %TODO
    }\label{fig:scenario-map}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
    %\includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
    \caption{
        Need a summary sentence.
        %TODO
        Trade-offs between (a) construction cost and expected lifetime costs or (b) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
    }
    \label{fig:tradeoffs}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics[width=3in]{lsl_priors.pdf}~
    %\includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
    \caption{
        The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
        Alternative priors can provide robustness checks.
        Need to add panel labels, update data, \etc.
        %TODO
    }\label{fig:tradeoffs}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics[width=6in]{inference_weights.pdf}
    \caption{
        The average weight assigned to \glspl{sow} from each model (\gls{rcp} scenario and ice sheet parameterization) for both priors considered.
        If the sampling distribution (green line) is used, then the weight assigned to each \gls{sow} is the same, regardless of whether it is over- or under-sampled.
    }\label{fig:model-average-weights}
\end{figure}

@@

\printbibliography

\appendix
\newcommand{\hbAppendixPrefix}{S}
\renewcommand{\thefigure}{\hbAppendixPrefix\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{\hbAppendixPrefix\arabic{table}}
\setcounter{table}{0}
\renewcommand{\theequation}{\hbAppendixPrefix\arabic{equation}}
\setcounter{equation}{0}

\newpage
\section{Supplemental Figures}
\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-depth-damage}
    \caption{
        Depth-damage relationship.
        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA.
        Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
    }\label{fig:cost-depth-damage}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-expected-damage-emulator}
    \caption{
        As discussed in XXX, we model expected annual damages as a function of the house's elevation relative to \gls{lsl}. % TODO: ref equation
    }\label{fig:cost-expected-damage-emulator}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-up-front}
    \caption{
        Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial fee.
        Values are sensitive to house floor area and structural value; see \cref{tab:uncertainties}.
    }\label{fig:cost-up-front}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{grid-sketch}
    \caption{
        Schematic of equation XXX used to calculate weights for each sample. % TODO: reference equation
        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
        As discussed in section XXX, the weights (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution  (horizontal dotted lines) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines). % TODO: reference section
    }\label{fig:grid-sketch}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
    }\label{fig:lsl-evolution}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-priors-weights}
    \caption{
        Subjective priors for local sea level.
        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
        The \glspl{pdf} of these subjective priors are shown in panel (A).
        In panels (B)-(D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
        Specifically, (B)-(D) show the average weight given to each model by each of the three subjective priors.
        The ``Slow SLR'' prior places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
        The ``Fast SLR'' prior places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
        The ``Intermediate SLR'' scenario places approximately equal weight on all models.
    }\label{fig:lsl-priors-weights}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{scenario-map-height-slr}
    \caption{
        Expected lifetime total cost (eq.XXX) as a function of height increase $\Delta h$ and sea level rise over the house lifetime. % TODO: reference equation
        The best outcomes (yellow) arise when the house is not elevated and sea level rise is minimal (bottom left corner).
        The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
        In general, a height increase of $<\SI{3}{ft}$ can be considered unrealistic \citep{zarekarizi_suboptimal:2020}.
        Values are sensitive to model constants; see \cref{tab:uncertainties}.
    }\label{fig:scenario-map-height-slr}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{scenario-map-slr-cost}
    \caption{scenario-map-slr-cost}\label{fig:scenario-map-slr-cost}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{surge-gev-priors}
    \caption{surge-gev-priors}\label{fig:surge-gev-priors}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{surge-obs-return}\label{fig:surge-obs-return}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-posterior-chains}
    \caption{surge-posterior-chains}\label{fig:surge-posterior-chains}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-prior-chains}
    \caption{surge-prior-chains}\label{fig:surge-prior-chains}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{surge-synthetic-data-experiment}
    \caption{surge-synthetic-data-experiment}\label{fig:surge-synthetic-data-experiment}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-test-statistics}
    \caption{surge-test-statistics}\label{fig:surge-test-statistics}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
    \caption{tradeoffs-by-prior}\label{fig:tradeoffs-by-prior}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
    \caption{tradeoffs-by-rcp}\label{fig:tradeoffs-by-rcp}
\end{figure}

\end{document}
